---
title: "Designing and Creating a Database"
author: "Dennis Jonathan"
date: "5/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We will work with data from [Major League Baseball](https://en.wikipedia.org/wiki/Major_League_Baseball) games compiled by [Retrosheet](http://www.retrosheet.org/), a non-profit organization that's gathered game statistics going back to the 1800s to today. The main file we will work from is [`game_log.csv`](https://dsserver-prod-resources-1.s3.amazonaws.com/376/game_log.csv), which has been compiled and pre-cleaned from 127 separate CSV files from Retrosheet. This file has hundreds of data points on each game. The goal of this mission is to convert and normalize this data into several separate tables using SQL and create a robust database of game-level statistics.

In addition to the main file, we also have four "helper" files sourced from Retrosheet:

- [`park_codes.csv`](https://dsserver-prod-resources-1.s3.amazonaws.com/376/park_codes.csv)
- [`person_codes.csv`](https://dsserver-prod-resources-1.s3.amazonaws.com/376/person_codes.csv)
- [`team_codes.csv`](https://dsserver-prod-resources-1.s3.amazonaws.com/376/person_codes.csv)
- [`appearance_type.csv`](https://dsserver-prod-resources-1.s3.amazonaws.com/376/appearance_type.csv)

These three helper files will make things easier since they will form the basis for three of our normalized tables.

## Prerequisites
```{r prereqs}
# Shutting off warning messages
options(warn=-1)

# Importing the libraries
library(tidyverse)
library(DBI)
library(RSQLite)
```

## Initial Data Exploration
We will now take our time to load the four CSV files we will be working with , look at its dimension and find out what the data is telling us.

### 1. Game Log

``` {r ide1}
# Reading the data in
log <- read_csv('game_log.csv', col_types = cols(.default = "c"))

# Checking the data dimension
dim(log)

# Previewing the first 5 data
head(log, 5)
```
It looks like the game log has a record of 171,907 games with 161 metrics on record ordered chronologically

For each game we have:

* general information on the game
* team level stats for each team
* a list of players from each team, numbered, with their defensive positions
* the umpires that officiated the game
* some 'awards', like winning and losing pitcher

We have a game_log_fields.txt file that tell us that the player number corresponds with the order in which they batted.

It's worth noting that there is no natural primary key column for this table. Also we have some columns which contains plenty of NAs.

### 2. Park Codes
```{r ide2}
# Importing the data in
park <- read_csv('park_codes.csv')

# Getting the data dimension
dim(park)

# Previewing the first five data
head(park, 5)
```
`park_codes.csv` contains the some information such as the name of the stadium/park, the park's id in the game log, the location of the stadium, and the league.

### 3. Person Codes
``` {r ide3}
# Importing the data in
person <- read_csv('person_codes.csv')

# Getting the data dimension
dim(person)

# Previewing the first five data
head(person, 5)
```
Similar to `park_codes.csv`, `person_codes.csv` contains the id for the players, their first and last name, as well as when they debuted as one of the roles in baseball.

### 4. Team Codes
``` {r ide4}
# Importing the data in
team <- read_csv('team_codes.csv')

# Getting the data dimension
dim(team)

# Previewing the first five data
head(team, 5)
```
`team_codes.csv` contains the team's ids, the league they are or were playing in, the city they are or were lockated and their starting and ending date.

It is quite clear that all of the ids available will be connected in one way or another to the game log later on when we are creating the database.

### 4. Defensive Positions

In the game log, each player has a defensive position listed, which seems to be a number between 1-10. Doing some research around this, I found this [article] (http://probaseballinsider.com/baseball-instruction/baseball-basics/baseball-basics-positions/) which gives us a list of names for each numbered position:

* Pitcher
* Catcher
* 1st Base
* 2nd Base
* 3rd Base
* Shortstop
* Left Field
* Center Field
* Right Field

The 10th position isn't included, it may be a way of describing a designated hitter that does not field. I can find a retrosheet page that indicates that position 0 is used for this, but we don't have any position 0 in our data. I have chosen to make this an 'Unknown Position' so I'm not including data based on a hunch.

### 5. Leagues

Wikipedia tells us there are currently two leagues - the American (AL) and National (NL). Upon investigation of the data, we see that there are actually 4 more. After some googling, we come up with:

* NL: National League
* AL: American League
* AA: [American Association](https://en.wikipedia.org/wiki/American_Association_%2819th_century%29)
* FL: [Federal League](https://en.wikipedia.org/wiki/Federal_League)
* PL: [Players League](https://en.wikipedia.org/wiki/Players%27_League)
* UA: [Union Association](https://en.wikipedia.org/wiki/Union_Association)

It also looks like we have about 1000 games where the home team doesn't have a value for league.

## Importing Data Into SQLite
We will now begin the process of creating our dataset, first and foremost we will create a `.db` file called `mlb.db` as the main framework to store everything. After that, we will create tables in there corresponding to the CSVs we have loaded in R previously.
```{r sqlite_import1}
# Creating the sql file
conn <- dbConnect(SQLite(), "mlb.db")

# Adding the csv into the sql
dbWriteTable(conn = conn, name = "game_log", 
             value = log, row.names = FALSE, header = TRUE)
dbWriteTable(conn = conn, name = "person_codes", 
             value = person, row.names = FALSE, header = TRUE)
dbWriteTable(conn = conn, name = "team_codes", 
             value = team, row.names = FALSE, header = TRUE)
dbWriteTable(conn = conn, name = "park_codes", 
             value = park, row.names = FALSE, header = TRUE)

# Checking whether we have successfully done our task
dbListTables(conn)
```
We have successfully exported the CSV files into our database, next up is to create another column in `game_log` which will be our primary key later on called `game_id`. That column will be the result of the concatenation of `date`, `h_name`, and `number_of_game`.
```{r sqlite_import2}
# Creating the game_id column
alter_game_log_command <- "
  ALTER TABLE game_log
  ADD COLUMN game_id TEXT;
"
dbExecute(conn, alter_game_log_command)

# Filling the game_id column
update_game_log_command <- "
  UPDATE game_log
  SET game_id = date || h_name || number_of_game
  WHERE game_id IS NULL;
"
dbExecute(conn, update_game_log_command)

# Checking the changes
check_game_id <- "
SELECT game_id,
       date,
       h_name,
       number_of_game
FROM game_log
LIMIT 3;
"
check <- dbGetQuery(conn, check_game_id)
head(check)
```
## Looking for Normalization Opportunities

We will now look to normalize the data that we have. In general we are looking for data which are redundant and data which are repetitive. After looking around, the following are opportunities for normalization of our data:

* In person_codes, all the debut dates will be able to be reproduced using game log data.
* In team_codes, the start, end and sequence columns will be able to be reproduced using game log data.
* In park_codes, the start and end years will be able to be reproduced using game log data. While technically the state is an attribute of the city, we might not want to have a an incomplete city/state table so we will leave this in.
* There are lots of places in game log where we have a player ID followed by the players name. We will be able to remove this and use the name data in person_codes
* In game_log, all offensive and defensive stats are repeated for the home team and the visiting team. We could break these out and have a table that lists each game twice, one for each team, and cut out this column repetition.
* Similarly, in game_log, we have a listing for 9 players on each team with their positions - we can remove these and have one table that tracks player appearances and their positions.
* We can do a similar thing with the umpires from game_log, instead of listing all four positions as columns, we can put the umpires either in their own table or make one table for players, umpires and managers.
* We have several awards in game_log like winning pitcher and losing pitcher. We can either break these out into their own table, have a table for awards, or combine the awards in with general appearances like the players and umpires.

We will then create our database using the following schema
![Alt text](https://dq-content.s3.amazonaws.com/376/mlb_schema.svg "MLB Schema")

## Creating Tables Which Do Not Have Foreign Keys
The title is self-explanatory. Essentially we are building a stable foundation for our database which will enable us to link the tables to other tables using foreign keys.

### 1. Person
```{r person}
# Creating the person table
create_person_command <- "
  CREATE TABLE IF NOT EXISTS person (
      person_id TEXT PRIMARY KEY,
      first_name TEXT,
      last_name TEXT
  );
"
dbExecute(conn, create_person_command)

# Filling the data to person from person_codes
insert_to_person <- "
  INSERT OR IGNORE INTO person
  SELECT
      id,
      first,
      last
  FROM person_codes;
"
dbExecute(conn, insert_to_person)

# Checking the query
check_query_2 <- "SELECT * FROM person LIMIT 5;"
check <- dbGetQuery(conn, check_query_2)
head(check)
```
### 2. Park
```{r park}
# Creating the park table
create_park_command <- "
  CREATE TABLE IF NOT EXISTS park (
      park_id TEXT PRIMARY KEY,
      name TEXT,
      nickname TEXT,
      city TEXT,
      state TEXT,
      notes TEXT
  );
"
dbExecute(conn, create_park_command)

# FIlling the park table
insert_to_park <- "
  INSERT OR IGNORE INTO park
  SELECT
      park_id,
      name,
      aka,
      city,
      state,
      notes
  FROM park_codes;
"
dbExecute(conn, insert_to_park)

# Checking the result
check_query_3 <- "SELECT * FROM park LIMIT 5;"
check <- dbGetQuery(conn, check_query_3)
head(check)
```
### 3. League
```{r league}
# Creating the league table
create_league_command <- "
  CREATE TABLE IF NOT EXISTS league (
      league_id TEXT PRIMARY KEY,
      name TEXT
  );
"
dbExecute(conn, create_league_command)

# Filling in the league table
insert_to_league <- '
  INSERT OR IGNORE INTO league
  VALUES
      ("NL", "National League"),
      ("AL", "American League"),
      ("AA", "American Association"),
      ("FL", "Federal League"),
      ("PL", "Players League"),
      ("UA", "Union Association")
  ;
'
dbExecute(conn, insert_to_league)

# Checking the result
check_query_4 <- "SELECT * FROM league"
check <- dbGetQuery(conn, check_query_4)
head(check)
```
### 4. Appearance type
```{r app_type}
# Dropping the table if it already exists
appearance_drop_precaution <- "DROP TABLE IF EXISTS appearance_type;"
dbExecute(conn, appearance_drop_precaution)

# Importing the csv into a table
appearance_type = read_csv('appearance_type.csv')
dbWriteTable(conn = conn, name = "appearance_type", 
             value = appearance_type,
             row.names = FALSE, header = TRUE)

# Checking the result
check_query_5 <- "SELECT * FROM appearance_type;"
check <- dbGetQuery(conn, check_query_5)
head(check)
```
## Tables with Relation to Other Tables
After finishing the building blocks, we can now start linking the tables we have just made and create the rest of the tables.

### 1. Team
```{r team}
# Creating the team table
create_team_command <- "
  CREATE TABLE IF NOT EXISTS team (
      team_id TEXT PRIMARY KEY,
      league_id TEXT,
      city TEXT,
      nickname TEXT,
      franch_id TEXT,
      FOREIGN KEY (league_id) REFERENCES league(league_id)
  );
"
dbExecute(conn, create_team_command)

# Filling the team table using data from team_codes
insert_to_team <- "
  INSERT OR IGNORE INTO team
  SELECT
      team_id,
      league,
      city,
      nickname,
      franch_id
  FROM team_codes;
"
dbExecute(conn, insert_to_team)

# Checking the result
check_query_6 <- "SELECT * FROM team LIMIT 5;"
check <- dbGetQuery(conn, check_query_6)
head(check)
```
### 2. Game
```{r game}
# Creating the game table
create_game_command <- "
  CREATE TABLE IF NOT EXISTS game (
      game_id TEXT PRIMARY KEY,
      date TEXT,
      number_of_game INTEGER,
      park_id TEXT,
      length_outs INTEGER,
      day BOOLEAN,
      completion TEXT,
      forfeit TEXT,
      protest TEXT,
      attendance INTEGER,
      legnth_minutes INTEGER,
      additional_info TEXT,
      acquisition_info TEXT,
      FOREIGN KEY (park_id) REFERENCES park(park_id)
  );
"

# Filling in the game table from game_log
dbExecute(conn, create_game_command)
insert_to_game <- '
  INSERT OR IGNORE INTO game
  SELECT
      game_id,
      date,
      number_of_game,
      park_id,
      length_outs,
      CASE
          WHEN day_night = "D" THEN 1
          WHEN day_night = "N" THEN 0
          ELSE NULL
          END
          AS day,
      completion,
      forfeit,
      protest,
      attendance,
      length_minutes,
      additional_info,
      acquisition_info
  FROM game_log;
'
dbExecute(conn, insert_to_game)

# Checking the result
check_query_7 <- "SELECT * FROM game LIMIT 5;"
check <- dbGetQuery(conn, check_query_7)
head(check)
```
### 3. Team appearance

```{r team_app}
# Creating the team appearance table
create_team_appearance_command <- "
  CREATE TABLE IF NOT EXISTS team_appearance (
      team_id TEXT,
      game_id TEXT,
      home BOOLEAN,
      league_id TEXT,
      score INTEGER,
      line_score TEXT,
      at_bats INTEGER,
      hits INTEGER,
      doubles INTEGER,
      triples INTEGER,
      homeruns INTEGER,
      rbi INTEGER,
      sacrifice_hits INTEGER,
      sacrifice_flies INTEGER,
      hit_by_pitch INTEGER,
      walks INTEGER,
      intentional_walks INTEGER,
      strikeouts INTEGER,
      stolen_bases INTEGER,
      caught_stealing INTEGER,
      grounded_into_double INTEGER,
      first_catcher_interference INTEGER,
      left_on_base INTEGER,
      pitchers_used INTEGER,
      individual_earned_runs INTEGER,
      team_earned_runs INTEGER,
      wild_pitches INTEGER,
      balks INTEGER,
      putouts INTEGER,
      assists INTEGER,
      errors INTEGER,
      passed_balls INTEGER,
      double_plays INTEGER,
      triple_plays INTEGER,
      PRIMARY KEY (team_id, game_id),
      FOREIGN KEY (team_id) REFERENCES team(team_id),
      FOREIGN KEY (game_id) REFERENCES game(game_id),
      FOREIGN KEY (team_id) REFERENCES team(team_id)
  );
"

# Filling the team appearance table
dbExecute(conn, create_team_appearance_command)
insert_to_team_appearance <- "
  INSERT OR IGNORE INTO team_appearance
      SELECT
          h_name,
          game_id,
          1 AS home,
          h_league,
          h_score,
          h_line_score,
          h_at_bats,
          h_hits,
          h_doubles,
          h_triples,
          h_homeruns,
          h_rbi,
          h_sacrifice_hits,
          h_sacrifice_flies,
          h_hit_by_pitch,
          h_walks,
          h_intentional_walks,
          h_strikeouts,
          h_stolen_bases,
          h_caught_stealing,
          h_grounded_into_double,
          h_first_catcher_interference,
          h_left_on_base,
          h_pitchers_used,
          h_individual_earned_runs,
          h_team_earned_runs,
          h_wild_pitches,
          h_balks,
          h_putouts,
          h_assists,
          h_errors,
          h_passed_balls,
          h_double_plays,
          h_triple_plays
      FROM game_log
  
  UNION
  
      SELECT    
          v_name,
          game_id,
          0 AS home,
          v_league,
          v_score,
          v_line_score,
          v_at_bats,
          v_hits,
          v_doubles,
          v_triples,
          v_homeruns,
          v_rbi,
          v_sacrifice_hits,
          v_sacrifice_flies,
          v_hit_by_pitch,
          v_walks,
          v_intentional_walks,
          v_strikeouts,
          v_stolen_bases,
          v_caught_stealing,
          v_grounded_into_double,
          v_first_catcher_interference,
          v_left_on_base,
          v_pitchers_used,
          v_individual_earned_runs,
          v_team_earned_runs,
          v_wild_pitches,
          v_balks,
          v_putouts,
          v_assists,
          v_errors,
          v_passed_balls,
          v_double_plays,
          v_triple_plays
      from game_log;
"
dbExecute(conn, insert_to_team_appearance)

# Checking the changes
check_query_8 <- "
  SELECT * FROM team_appearance
  WHERE game_id = (
                   SELECT MIN(game_id) from game
                  )
     OR game_id = (
                   SELECT MAX(game_id) from game
                  )
  ORDER By game_id, home;
"
check <- dbGetQuery(conn, check_query_8)
head(check)
```
### 4. Person appearance
```{r person_app}
# Creating the person appearance table
create_person_appearance_command <- "
  CREATE TABLE person_appearance (
      appearance_id INTEGER PRIMARY KEY,
      person_id TEXT,
      team_id TEXT,
      game_id TEXT,
      appearance_type_id,
      FOREIGN KEY (person_id) REFERENCES person(person_id),
      FOREIGN KEY (team_id) REFERENCES team(team_id),
      FOREIGN KEY (game_id) REFERENCES game(game_id),
      FOREIGN KEY (appearance_type_id) REFERENCES appearance_type(appearance_type_id)
  );
"
dbExecute(conn, create_person_appearance_command)

# Inputing the data into person appearance
insert_to_person_appearance <- '
  INSERT OR IGNORE INTO person_appearance (
      game_id,
      team_id,
      person_id,
      appearance_type_id
  ) 
      SELECT
          game_id,
          NULL,
          hp_umpire_id,
          "UHP"
      FROM game_log
      WHERE hp_umpire_id IS NOT NULL    
  
  UNION
  
      SELECT
          game_id,
          NULL,
          [1b_umpire_id],
          "U1B"
      FROM game_log
      WHERE "1b_umpire_id" IS NOT NULL
  
  UNION
  
      SELECT
          game_id,
          NULL,
          [2b_umpire_id],
          "U2B"
      FROM game_log
      WHERE [2b_umpire_id] IS NOT NULL
  
  UNION
  
      SELECT
          game_id,
          NULL,
          [3b_umpire_id],
          "U3B"
      FROM game_log
      WHERE [3b_umpire_id] IS NOT NULL
  
  UNION
  
      SELECT
          game_id,
          NULL,
          lf_umpire_id,
          "ULF"
      FROM game_log
      WHERE lf_umpire_id IS NOT NULL
  
  UNION
  
      SELECT
          game_id,
          NULL,
          rf_umpire_id,
          "URF"
      FROM game_log
      WHERE rf_umpire_id IS NOT NULL
  
  UNION
  
      SELECT
          game_id,
          v_name,
          v_manager_id,
          "MM"
      FROM game_log
      WHERE v_manager_id IS NOT NULL
  
  UNION
  
      SELECT
          game_id,
          h_name,
          h_manager_id,
          "MM"
      FROM game_log
      WHERE h_manager_id IS NOT NULL
  
  UNION
  
      SELECT
          game_id,
          CASE
              WHEN h_score > v_score THEN h_name
              ELSE v_name
              END,
          winning_pitcher_id,
          "AWP"
      FROM game_log
      WHERE winning_pitcher_id IS NOT NULL
  
  UNION
  
      SELECT
          game_id,
          CASE
              WHEN h_score < v_score THEN h_name
              ELSE v_name
              END,
          losing_pitcher_id,
          "ALP"
      FROM game_log
      WHERE losing_pitcher_id IS NOT NULL
  
  UNION
  
      SELECT
          game_id,
          CASE
              WHEN h_score > v_score THEN h_name
              ELSE v_name
              END,
          saving_pitcher_id,
          "ASP"
      FROM game_log
      WHERE saving_pitcher_id IS NOT NULL
  
  UNION
  
      SELECT
          game_id,
          CASE
              WHEN h_score > v_score THEN h_name
              ELSE v_name
              END,
          winning_rbi_batter_id,
          "AWB"
      FROM game_log
      WHERE winning_rbi_batter_id IS NOT NULL
  
  UNION
  
      SELECT
          game_id,
          v_name,
          v_starting_pitcher_id,
          "PSP"
      FROM game_log
      WHERE v_starting_pitcher_id IS NOT NULL
  
  UNION
  
      SELECT
          game_id,
          h_name,
          h_starting_pitcher_id,
          "PSP"
      FROM game_log
      WHERE h_starting_pitcher_id IS NOT NULL;
'
dbExecute(conn, insert_to_person_appearance)

# Inputing offensive and defensive player data into person appearance
for (letter in c("h", "v")) {
  for (num in 1:9) {
    template <- '
      INSERT INTO person_appearance (
          game_id,
          team_id,
          person_id,
          appearance_type_id
      ) 
          SELECT
              game_id,
              %s_name,
              %s_player_%f_id,
              "O%f"
          FROM game_log
          WHERE %s_player_%f_id IS NOT NULL

      UNION

          SELECT
              game_id,
              %s_name,
              %s_player_%f_id,
              "D" || CAST(%s_player_%f_def_pos AS INT)
          FROM game_log
          WHERE %s_player_%f_id IS NOT NULL;
    '
    # replace all of the %s and %f with the correct letter number
    template <- gsub("%s", letter, template, fixed = TRUE)
    template <- gsub("%f", num, template, fixed = TRUE)
  }
}
```
## Deleting the Old Tables
Now that we have finished creating our database, we can delete the old database since those are no longer useful for us.

```{r drop}
# Check the current status of the db file
dbListTables(conn)

# Iterate through each of the tables
tables <- c("game_log", "park_codes",
            "team_codes", "person_codes")
for (t in tables) {
  drop_command = sprintf("DROP TABLE %s", t)
  dbExecute(conn, drop_command)
}
# Make sure that everything is gone
dbListTables(conn)

# Stopping the connection
dbDisconnect(conn)
```
